%tel% LaTeX Template für Abgaben an der Universität Stuttgart
% Autor: Sandro Speth
% Bei Fragen: Sandro.Speth@studi.informatik.uni-stuttgart.de
%-----------------------------------------------------------
% Hauptmodul des Templates: Hier können andere Dateien eingebunden werden
% oder Inhalte direkt rein geschrieben werden.
% Kompiliere dieses Modul um eine PDF zu erzeugen.

% Dokumentenart. Ersetze 12pt, falls die Schriftgröße anzupassen ist.
\documentclass[12pt,ngerman]{scrartcl}
\usepackage{superiors}
\usepackage{amsmath}
% Einbinden der Pakete, des Headers und der Formatierung.
% Mit den \include und \input Befehlen können Dateien eingebunden werden:
% \include: Fügt einen Seitenumbruch nach dem Text ein
% \input: Fügt KEINEN Seitenumbruch nach dem Text ein
\input{../styles/Packages.tex}
\input{../styles/FormatAndHeader.tex}

% Counter für das Blatt und die Aufgabennummer.
% Ersetze die Nummer des Übungsblattes und die Nummer der Aufgabe
% den Anforderungen entsprechend.
% Definiert werden die Counter in FormatAndHeader.tex
% Beachte:
% \setcounter{countername}{number}: Legt den Wert des Counters fest
% \stepcounter{countername}: Erhöht den Wert des Counters um 1.
\setcounter{sheetnr}{2} % Nummer des Übungsblattes
\setcounter{exnum}{0} % Nummer der Aufgabe

\def\domsum{\sum_{z\in\text{Dom}(Z)}}
% Beginn des eigentlichen Dokuments

\lstset{breaklines=true,
        breakatwhitespace=false,
        postbreak=\mbox{$\hookrightarrow$\space},
        basicstyle=\ttfamily\footnotesize,
        columns=fullflexible}

\begin{document}

\exercise{XML Schema}
\subexercise{}
A well-formed XML document must follow these two rules: 
\begin{itemize}
    \item First, it has to have exactly one top-level (root) element that contains all other elements.
    \item Second, it has to be nested properly. This means every start tag \texttt{<tag>} must have a matching closing tag \texttt{</tag>}, and different tags must not overlap (e.g. not allowed: \texttt{<a><b></a></b>}).
\end{itemize}
In this case the document has just one top-level element \texttt{<tns:Orders>}. It is also properly nested, as every tag has a matching closing tag and no tags overlap. The document is therefore well-formed.

\subexercise{}
\begin{itemize}
    \item Lines 17--21: The DeliveryAddress has the type tAddress, which requires the element \texttt{<StreetNumber>}. Since the elements of an \texttt{xs:all} group have a default minOccurs of 1, StreetNumber is mandatory but missing here.
    \item Line 26: The Quantity is 10, but the schema restricts Quantity with maxExclusive value 10, so the maximum allowed value is 9.
    \item Line 41: The BillingAddress also has the type tAddress, so it is required to have an addressID.
    \item Line 44: The ZipCode of a tAddress type has to consist of exactly five digits between 0 and 9. This one has only four.
    \item Lines 48--52: The child elements of the Item are in the wrong order. The type tOrderItem is defined as an \texttt{xs:sequence} (Name, Size, Quantity), but here Quantity appears first.
    \item Line 53: The itemID has to be unique, but the ID I000328 has already been used in line 48.
\end{itemize}

\subexercise{}
    No, one of the criteria for a document to be valid is, that it is well-formed.

\exercise{XQuery}

\subexercise{}
{\footnotesize
\begin{lstlisting}
for $hotel in doc("hotels.xml")//Hotel
return <output>{ concat($hotel/Name, " ", $hotel/Address/TownName, " ",
                        count($hotel//Reservation)) }</output>
\end{lstlisting}

\subexercise{}
\begin{lstlisting}
for $hotel in doc("hotels.xml")//Hotel,
    $guest in $hotel/Guests/Guest,
    $res in $hotel//Reservation[@guestNumber = $guest/@guestNumber]
return <output>{ $guest/Name/text() } { $res }</output>
\end{lstlisting}

\subexercise{}
\begin{lstlisting}
for $hotel in doc("hotels.xml")//Hotel,
    $guest in $hotel/Guests/Guest
where string-length($guest/Name) < 11
return <output>{ concat($hotel/@hotelNumber, " ", $guest/Name) }</output>
\end{lstlisting}

\subexercise{}
\begin{lstlisting}
for $hotel in doc("hotels.xml")//Hotel,
    $res in $hotel//Reservation[RoomType/BasicRoom/MiniBar = "false"],
    $guest in $hotel/Guests/Guest[@guestNumber = $res/@guestNumber]
order by $guest/Address/StreetAndHouseNumber ascending
return <output>{ concat($guest/@guestNumber, " ",
                        $guest/Address/StreetAndHouseNumber) }</output>
\end{lstlisting}
}

\exercise{DTD and XML Schema}
\subexercise{}
The DTD:
\lstinputlisting{machines.dtd}
An example document that uses this DTD:
\lstinputlisting[language=Xml]{machines.xml}
There are some requirements, that a DTD-document can't fullfill:
\begin{itemize}
    \item A machine-number must have exactly 5 characters. The DTD-document can't restrict the length of the data.
    \item A machine-temperature must contain a floating point number. The DTD-document can't restrict text to floating point numbers.
    \item Using the ID type to enforce uniqueness of machine-number restricts the value to a valid XML Name, which must not start with a digit. A purely numeric 5-character machine-number is therefore not expressible in a DTD. Likewise, incident-date cannot be validated as a date, only as \#PCDATA.
\end{itemize}

\subexercise{}
\lstinputlisting[language=Xml]{machines.xsd}

\subexercise{}
\begin{itemize}
    \item One advantage is, that with a XML-schema, there is a bigger differenciation between simple data types. In DTD you only have \#PCDATA, while you have integer, strings, dates, ... in a XML-schema.
    \item Another advantage would be the wider range for an XML-schema to have more complex structures due to things like minOccurs, maxOccurs, ...
\end{itemize}

\end{document}
